Selective enhancement of an exploited species destabilizes ecological communities

### 2021-11-15 rmd/manuscript.Rmd
Update manuscript.Rmd

---
title: "Excessive release of captive-bred animals destabilizes ecological communities"
output:
  word_document:
    reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

We developed a multispecies Ricker model to simulate community dynamics with selective enhancement of a particular species. SUPPLEMENTAL EXPLANATION HERE. Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = [N_{i,t}+\phi R_t]exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i})]exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient (intraspecific if $i=j$, interspecific otherwise), $K_i$ the carrying capacity, $R_t$ the number of released individuals for enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species.

>>>>

---
title: "Selective enhancement of an exploited species destabilizes ecological communities"
output:
  word_document:
    reference_docx: word_format.docx
csl: pnas.csl
knit: (function(inputFile, encoding) {
        rmarkdown::render(inputFile, encoding = encoding, output_dir = "document_output")
      })
bibliography: reference.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(tidyverse)

# theory infomation
df_theory <- read_csv(here::here("code_theory/result/result_ricker.csv"))

param_set <- df_theory %>% 
  filter(r_min == 0.5) %>% 
  distinct(n_species,
           k,
           r_type,
           r_min,
           r_max,
           sd_env,
           phi,
           int_type,
           alpha,
           model,
           seed)

```

@@

We developed a multispecies Ricker model to simulate community dynamics with selective enhancement of a particular species. Population density of species $i$ at time $t+1$, $N_{i,t+1}$, is modeled as:

$$
N_{i,t+1} = [N_{i,t}+\phi_i R_t]exp[r_i(1-\frac{N_{i,t}+\alpha_{i1} R_t + \sum_{j=1}^S \alpha_{ij}N_{j,t}}{K_i})]exp(\epsilon_{i,t})
$$

where $r_i$ is the intrinsic growth rate, $\alpha_{ij}$ the competition coefficient between species $i$ and $j$ (intraspecific competition if $i=j$; $\alpha_{ii}=1$), $K_i$ the carrying capacity, $R_t$ the number of released individuals for the enhancement of species 1, and $\epsilon_{i,t}$ the species responses to stochastic environmental fluctuations. The parameter $\phi_i$ takes a value of one when $i=1$ (otherwise zero) to represent the reproductive contribution of released individuals to the next generation of the enhanced species. The species-specific intrinsic growth rates and competition coefficients were drawn randomly from a uniform and exponential distribution, respectively, to allow for interspecific variation as $r_i \sim Unif(0.5, r_{max})$ and $\alpha_{ij, i \ne j} \sim Exp(\mu_{\alpha})$. Environmental stochasticity is random and was expressed as $\epsilon_{i,t} \sim Normal(\mu_{\epsilon}, \sigma_{\epsilon}^2)$

We varied values of `r sum(unlist(lapply(sapply(param_set, unique), length)) > 1)` parameters: $S =$ `r unique(param_set$n_species)`; $r_{max}=$ `r unique(param_set$r_max)`; $\mu_{\alpha} =$ `r unique(param_set$alpha)`; $\sigma_\epsilon =$ `r unique(param_set$sd_env)`. For simplicity, we assumed identical carrying capacities among simulated species as $K_i =$ `r unique(param_set$k)`. This setup results in `r nrow(param_set)` sets of parameter combinations that cover a range of ecological scenarios.

Under each parameter combination, we ran 1400 generations to obtain the relative temporal fluctuations (i.e., the coefficient of variation) of the whole community ($\sum_i^SN_j$), the enhanced species ($N_1$), and the unenhanced species ($\sum_{i,i \ne 1}^S N_i$). We also calculated temporal means and standard deviations for each species group.
